---
title: "CW1 main file"
author: "Edward Macintyre"
date: "27/02/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(tidyverse)
library(ggpubr)
library(GGally)

data_raw <- read_csv("CW1_data.csv")

# We can drop the first column since it is identical to the state variable
data_raw <- data_raw[-1]

# All of the variable names are fairly self-explanatory. 
# The Gini Index (giniindex) is a measure of income distribution across a population. 0 indicates perfect equality, and 1 indicates perfect inequality. i.e. If each person in a population earned the same income, the Gini Index would be 0, but if 1 person earned all of the income of a population, the Gini Index would be 1. A higher Gini Index indicates higher earners in a population earn a greater proportion of the total income.

# Split data into data to be used in Q1 and Q2 respectively. Some of the socioeconomic variables provided are a mix of 2015/2016 data, however, since they are averages/proportions from a very large sample size, it is safe to assume the values will be approximately equal in both year. i.e. we can use the proportion of the population that is non-white in 2015 as an estimate for that of 2016. Similarly, the proportion of adults 25 and older with a high-school degree in 2009 can be used as an estimate for that of 2010.
vars_Q1 <- c("state", "householdincome", "unemployed", "metropop", "noncitizen", "whitepoverty", "giniindex", "nonwhite", "trumpvoter", "incidents2016", "pop2016", "hrate2016")
vars_Q2 <- c("state", "metropop")

# 1. Did the 2016 election campaigns of Donald Trump lead to an increase in hate crimes?

# Initial thoughts
# To answer this question, I propose constructing a model based off of the 2016 data. We want to relate hate crime incidence to the proportion of population that voted for trump. A good model may be a poisson model with log link and incidence as the explanatory variable. 

# Look at modelling the difference between 2015 and 2016 incidences

# Look at correlations between other explanatory variables



# 2. Using data of the period 2010 to 2015, investigate whether there are associations between hate crime
# incident rates and the explanatory variables on socio-economic factors. Describe the strength and the
# nature of the association.
```

Caveats (Please add to these as you go along so we can mention them in the report):

1. The hate crime incidence broken down by bias/motivation is not specified in this particular data set
2. The hate crime incidence is only based off of crimes that are reported, do not mistake this for total number of hate crimes since many may go unreported
3. The state list contains all of the states of America, except Hawaii. This is because "Hawaii does not participate in the Hate Crime Statistics Program" (FBI)
4. The state list includes the District of Columbia (Washington D.C.) which isn't technically a state, however it functions as a state for all of the purposes we will be looking in to
5. The data provided is not super specific to Q1 so make sure to mention this
