---
title: "CW1 Q1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
# Packages and import data
library(tidyverse)
library(ggpubr)
library(GGally)
library(ggplot2)
dat <- read.csv("CW1_data.csv", header=T)
dat <- as.tibble(dat)

# First column the same as the second so remove it
dat <- dat[-1]
```

## Initial Checks

```{r}
summary(dat)
```

Appears to be no factor variables given the initial set up of the data.


## Investigate NAs in noncitizen variable

```{r}
plot(dat$noncitizen)
```

Not exactly a great range of values since you would expect most of the population to be citizens. Are citizens just as likely to experience HC as non citizens given say both being of a certain race or gender?

Can't delete the 3 NA observations since would lose 3/50 states. Valid reason to exclude the variable from consideration? Is the variable useful anyway given the other variables?

Lean towards just not considering the variable in any models given the variety on show in the other possible variables eg. unemployed, nonwhite, whitepoverty etc.


## Pairs and Correlations

```{r}
vars <- colnames(dat)[-6]
vars <- vars[-1]
round(cor(dat[, vars]), 2)
```

```{r}
ggpairs(data=dat[,vars])
```

```{r}
par(mfrow=c(1,2))
plot(dat$trumpvoter, dat$hrate201015)
plot(dat$trumpvoter, dat$hrate2016)
```

Plots appear to show that rate of HC decreased slightly in 2016 compared to 2010-15 average. Correlation confirms this observation. Clear outlier value is Washington DC. Might have to look at excluding this State for the purposes of the model, HC rate wildly different to all other state (not a state anyway).


## Removing Washington DC

```{r}
# Filter out District of Columbia
dat_xDC <- dat %>% filter(state!="District of Columbia")
```

Looking at correlations again

```{r}
round(cor(dat_xDC[, vars]), 2)
```

```{r}
par(mfrow=c(1,2))
plot(dat_xDC$trumpvoter, dat_xDC$hrate201015)
plot(dat_xDC$trumpvoter, dat_xDC$hrate2016)
```

Removing DC lessens the correlation however it is still negative.


```{r}
par(mfrow=c(1,2))
plot(density(dat$hrate201015), ylim=c(0,0.35))
lines(density(dat$hrate2016), col="red")
plot(density(dat_xDC$hrate201015), ylim=c(0,0.35))
lines(density(dat_xDC$hrate2016), col="red")
```

Plots above again don't show any major increases/decreases in the rate of HC in 2016 compared to 2010-15. If anything it shows a slight decrease.


## Making a "difference" variable and initial investigation

```{r}
# Creating difference variables for investigation
dat_xDC <- mutate(dat_xDC, incidentsDIFF=incidents2016-incidents201015, popDIFF=pop2016-pop201015, hrateDIFF=hrate2016-hrate201015)
```

```{r}
summary(dat_xDC[c(17,18,19)])
par(mfrow=c(1,2))
plot(density(dat_xDC$incidentsDIFF))
plot(density(dat_xDC$hrateDIFF))
```

"Differences" appear to be centered around zero possible slight skew to left (decrease) on hrateDIFF.


## First model - hrateDIFF as response

```{r}
DIFFlm_full <- lm(hrateDIFF ~ householdincome + unemployed + metropop + highschoolpop + whitepoverty + giniindex + nonwhite + trumpvoter, data=dat_xDC)
summary(DIFFlm_full)
```

Linear model since poisson can't take negative values and some differences are obviously negative.

```{r}
summary(DIFFlm_step <- step(DIFFlm_full, trace=0))
```


```{r}
plot(density(residuals(DIFFlm_step)))

par(mfrow=c(1,2))
plot(DIFFlm_step,c(1,2))

par(mfrow=c(2,2))
plot(dat_xDC$householdincome, residuals(DIFFlm_step))
plot(dat_xDC$metropop, residuals(DIFFlm_step))
plot(dat_xDC$trumpvoter, residuals(DIFFlm_step))
```

Doesn't seem to good, poor residual diagnostics, no clear indications of how to improve model. Rsquared low.
Conclusions not clear?

Maybe modelling DIFF as response isn't the appropriate method.




## Model 2016 rate as response


Since the response is rate it is not an integer, therefore cannot use poisson despite rate being derived from count. Instead use a Gamma.


```{r}
rateGLM_full <- glm(hrate2016 ~ householdincome + unemployed + metropop + highschoolpop + whitepoverty + giniindex + nonwhite + trumpvoter, data=dat_xDC, family=Gamma(log))
summary(rateGLM_step <- step(rateGLM_full, trace=0))
```


```{r}
plot(density(residuals(rateGLM_step)))

par(mfrow=c(1,2))
plot(rateGLM_step,c(1,2))
```

Residual diagnostics are okay.
Model suggests:
Higher white poverty percentage leads to higher HC rate.
Higher nonwhite populations (percentage) results in lower HC rate.
Higher proportion of trump voters leads to lower HC rate.



Try using incident count as response with population as an offset:

```{r}
incidentGLM_full <- glm(log(incidents2016) ~ householdincome + unemployed + metropop + highschoolpop + whitepoverty + giniindex + nonwhite + trumpvoter, offset=log(pop2016), data=dat_xDC, family=Gamma(log))
summary(incidentGLM_step <- step(incidentGLM_full, trace=0))
```

"Algorithm did not converge" ????




DO WE NEED TO MAKE A MODEL FOR THIS FIRST QUESTION????

SEEMS DATA CLEARLY SHOWS THAT HATE CRIME RATES DECREASED FROM 2010-15 TO 2016, CERTAINLY DIDN'T INCREASE. SO HOW CAN WE APPLY ANY OF THESE CONCLUSIONS TO TRUMP CAMPAIGNS????

ONLY LINK TO TRUMP IN DATA APPEARS TO BE TRUMP VOTER PERCENTAGE (UNLESS WE MAKE SWEEPING ASSUMPTIONS ABOUT CHARACTERISTICS OF TRUMP VOTERS) AND HIGHER PROPORTION OF TRUMP VOTERS APPEAR TO LEAD TO LOWER HATE CRIME RATES - CONCLUSIONS????

MAYBE HAVE THE WRONG APPROACH????



## Fitting a linear model with rate as response despite rate being strictly positive

```{r}
rateLM_full <- lm(hrate2016 ~ householdincome + unemployed + metropop + highschoolpop + whitepoverty + giniindex + nonwhite + trumpvoter, data=dat_xDC)
summary(rateLM_step <- step(rateLM_full, trace=0))
```

```{r}
plot(density(residuals(rateLM_step)))

par(mfrow=c(1,2))
plot(rateLM_step,c(1,2))

par(mfrow=c(2,2))
plot(dat_xDC$highschoolpop, residuals(rateLM_step))
plot(dat_xDC$whitepoverty, residuals(rateLM_step))
plot(dat_xDC$trumpvoter, residuals(rateLM_step))
```

Diagnostics seem okay, not great mind. Residual summary not perfect, low Rsquared.

Model suggests:
Higher percentage of population reaching the high school level, the higher hate crime rate.
Higher white poverty rates increase hate crime rate.
Higher trump voter percentage lowers the hate crime rate.

WHO KNOWS WHATS GOING ON

```{r}
AIC(rateLM_step)
AIC(rateGLM_step)
```
AIC prefers Gamma GLM.




WHO KNOWS HOW TO USE ANY OF THESE MODELS TO ANSWER THE QUESTION. QUESTION SEEMS TO GENERAL FOR THE DATA AND VARIABLES GIVEN.
ONLY VARIABLES RELATING TO TRUMP ELECTION IS TIME PERIOD, AND TRUMP VOTER PERCENTAGE.

EVEN WITHOUT MAKING THE MODELS, RATE APPEARS TO DECREASE SLIGHTLY, AND REASONABLE NEGATIVE CORRELATION WITH TRUMP VOTER AND 2016 RATE.

CAN'T CONFIDENTLY CONCLUDE (AT THE MOMENT) THAT RATE INCREASED POST ELECTION, LET ALONE DUE TO THE TRUMP CAMPAIGNS. IF ANYTHING, MODELS AND EXPLANATORY ANALYSIS INDICATE A DECREASE WHICH SEEMS UNUSUAL (GIVEN PRE ASSUMED THOUGHTS ON TRUMP PRESIDENCY).

NO EVIDENCE TO SUPPORT STATEMENT IN QUESTION 1!!!!!!!





## LIMITATIONS:

IDEALLY WOULD CONDUCT TIME SERIES ANALYSIS FOR A BETTER UNDERSTANDING OF ANY TREND

SOME HATE CRIMES WILL NOT BE REPORTED LEADING TO INCOMPLETE DATA

WOULD BE GOOD TO HAVE INDIVIDUAL HATE CRIMES SPLIT OUT IE. RACISM, SEXISM ETC ETC TO SEE IF THERE IS A TREND FOR SPECIFIC HATE CRIME, IE RACISM INCREASED BUT SEXISM DECREASED, THAT EFFECT WOULD BE MASKED IN THIS DATA SET.

ONLY 50 STATES (49 ONCE DC REMOVED) SO SMALL DATA SET, MAYBE A LARGER DATA SET COULD SPLIT STATES FURTHER INTO COUNTIES SAY, HOWEVER THIS WOULD THEN BE AFFECTED BY LOW REPORTED NUMBERS, IE WOULD EXPECT MOUNTAINS IN MONTANA TO HAVE VERY FEW REPORTED HATE CRIMES.








